---
title: "Data Science Specialism"
output:
  html_document:
    df_print: paged
---

# Reading HDF5 Data (a hierarchical data format with groups and datasets)
https://www.coursera.org/learn/data-cleaning/home/welcome

<br />

HDF5's main features are the ability to store and access very large and/or complex datasets and a wide variety of metadata on mass storage (disk) through a completely portable file format. The rhdf5 package is thus suited for the exchange of large and/or complex datasets between R and other software package, and for letting R applications work on datasets that are larger than the available RAM.


**Further Info**

HDF5: https://hdf5group.org

Docs: https://bioconductor.org/packages/release/bioc/html/rhdf5.html

Tutorial: https://www.neonscience.org/resources/learning-hub/tutorials/hdf5-intro-r


<br />

```{r setup}
# {r setup}

mod3Dir <- "~/repositories/learning/data-science-coursera/03-get-clean-data"
dataDir <- "data"

setwd(mod3Dir)
if(!file.exists(dataDir)) { dir.create(dataDir) }

# set the root directory for this notebook
knitr::opts_knit$set(root.dir = paste(mod3Dir, dataDir, sep = "/"))

# delete the data file if it exists
setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")
if(file.exists("steven.h5")){
        file.remove("steven.h5")
}


# install the prerequisites
# https://bioconductor.org/packages/release/bioc/html/rhdf5.html
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("rhdf5")

library("rhdf5")
```


<br />

#### Creating an HDF5 file
```{r}
# setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")
created <- h5createFile("steven.h5")
created
```

<br />

#### Creating Groups and Datasets in the HDF5 file
```{r}
# setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")
# "created" just swallows the [1] TRUE output
created <- h5createGroup("steven.h5", "money")
created <- h5createGroup("steven.h5", "health")
created <- h5createGroup("steven.h5", "learning")
created <- h5createGroup("steven.h5", "learning/data-science")

h5ls("steven.h5")
```

```{r}
# writing data to a group
# create matrices and write them to groups
# setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")

accounts = matrix(1:10, nr=5,nc=2)
h5write(accounts, file="steven.h5", "/money/accounts")


# can also write data to the top level group
letts <- data.frame(1L:5L, seq(0,1,length.out = 5),
                    c("ab", "cde", "fghi", "a", "s"),
                    stringsAsFactors = FALSE
)
h5write(letts, "steven.h5", "letters")

h5ls("steven.h5")
```

```{r}
# reading the data from an HDF5 file
# setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")

readLetrs <- h5read("steven.h5", "letters")

readLetrs
```

```{r}
# writing to specific parts of a dataset
# setwd("~/repositories/learning/data-science-coursera/03-get-clean-data/data")


# the index argument can also be passed to h5read if you want to read specific parts
# of a dataset in and HDF5 document
h5write(c(12,13,14), "steven.h5", "money/accounts", index=list(1:3,1))
h5read("steven.h5", "money/accounts")
```
