---
title: "Data Science Specialism"
output:
  html_document:
    df_print: paged
---

# Getting and Cleaning Data

https://www.coursera.org/learn/data-cleaning/home/welcome

### Week 1

#### Downloading Files

Use the ```setup``` chunk to set the working directory for all other chunks in this notebook. If we did this in a normal chunk the working directory would be reset as soon as that chunk had finished executing.

```{r setup}
# {r setup}

mod3Dir <- "~/repositories/learning/data-science-coursera/03-get-clean-data"
dataDir <- "data"

setwd(mod3Dir)
if(!file.exists(dataDir)) { dir.create(dataDir) }

# set the root directory for this notebook
knitr::opts_knit$set(root.dir = paste(mod3Dir, dataDir, sep = "/"))
```


<br />

#### Question 1 of the the Quiz.

```{r}
# Question 1 - count the number of properties whose value in $1,000,000 or more

Q1dataUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
Q1dataDestFile <- "american_community_survey_2006.csv"

if(!file.exists(Q1dataDestFile)) {
        download.file(Q1dataUrl, Q1dataDestFile, method = "curl")
        Q1dataDownloaded <- date()
}

Q1codebookUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf"
Q1codebookDestFile <- "american_community_survey_code_book.pdf"

if(!file.exists(Q1codebookDestFile)) {
        download.file(Q1codebookUrl, Q1codebookDestFile, method = "curl")
        Q1codebookDownloaded <- date()
}

# Load the data
communityData <- read.csv(Q1dataDestFile)

# answer the question
length(which(communityData$VAL == 24))
```

<br />

#### Readng the downloaded data

Some choices

+ ```read.table(file, header, sep, row.names, nrows)``` - robust, flexible but slow 
+ ```read.csv(file, header = TRUE, sep = ",")``` - other options below

```quote = ""``` - the character to treat as a quote character in the CSV file - set to ```""``` if you are having trouble with data that has ``` ` ``` or ```"``` characters in the data

```na-strings = "NA"``` - the string used in the file to represent NA

```nrows = 100``` - the number of rows to read

```skip = 1``` - the number of rows to skip before starting to read the file.

<br />

#### Question 3 of the the Quiz.


```{r}
# Question 3 - 


Q3dataUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
Q3dataDestFile <- "naturalGasAcquisitionDestFile.xlsx"

if(!file.exists(Q3dataDestFile)) {
        download.file(Q3dataUrl, Q3dataDestFile, method = "curl")
        Q3dataDownloaded <- date()
}

# Load the data - read rows 18-23 and columns 7-15 and assign to "dat"

# to get this to work I had to install rJava
# (https://www.r-bloggers.com/2018/02/installing-rjava-on-ubuntu/)
# as this is a dependency of the xlsx package

# also to get the rJava package to work I had to do the following
# export LD_LIBRARY_PATH=/usr/lib/jvm/java-8-openjdk-amd64/jre/lib/amd64/server
# sudo R CMD javareconf


library(xlsx)
colIndex <- 7:15
rowIndex <- 18:23
dat <- read.xlsx(Q3dataDestFile, sheetIndex = 1, colIndex = colIndex, rowIndex = rowIndex)
# answer the question - what is the output of the following function
sum(dat$Zip*dat$Ext,na.rm=T)
```

<br />

#### Question 4 of the the Quiz.

Get extra info on extracting data from XML here:
https://www.stat.berkeley.edu/~statcur/Workshop2/Presentations/XML.pdf

```{r}
# Question 4 - 

# in order to get this to work I hadf to install the Ubuntu package libxml2-dev
# > sudo apt-get install libxml2-dev
# > R
# > install.packages("XML")


library(XML)

Q4dataUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
Q4dataDestFile <- "restaurants.xml"

if(!file.exists(Q4dataDestFile)) {
        download.file(Q4dataUrl, Q4dataDestFile, method = "curl")
        Q4dataDownloaded <- date()
}

# this did not work because xmlTreeParse does not support https
# Q4Doc <- xmlTreeParse(Q4dataUrl)

# so instead I had to do this:
Q4Doc <- xmlTreeParse(sub("s", "", Q4dataUrl), useInternal = TRUE)
rootNode <- xmlRoot(Q4Doc)

# answer the question - what is the output of the following function
# this by the way is not great at all
zips <- xmlSApply(rootNode[[1]], function(x) xmlSApply(x[["zipcode"]], xmlValue))
tableOfZips <- table(zips)
zip21231 <- tableOfZips[names(tableOfZips) == 21231]
zip21231
```


<br />

#### Question 5 of the the Quiz.


```{r}
# Question 5 - 



Q5dataUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
Q5dataDestFile <- "idaho.csv"

if(!file.exists(Q5dataDestFile)) {
        download.file(Q5dataUrl, Q5dataDestFile, method = "curl")
        Q5dataDownloaded <- date()
}



# need to install and load the package data.table in order to use fread()
# > install.packages("data.table")
library("data.table")

#fread() - fast read detects all setting
DT <- fread(Q5dataDestFile)



```

